Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-22 paper/results.tex
Clarified the conditions for the detailed timing tests. Added HADOOP line count. Other minor changes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@675 3fc257bb-8550-0410-81b3-a5c103da83f7

All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
0.05)$ and generated 8 edges per vertex (on average).  These
parameters create a highly skewed degree distribution, as indicated by
the maximum vertex degree in Table~\ref{t:rmats}.

@@

A more detailed analysis of the performance of the MR-MPI PageRank
algorithm is shown below.  We perform the analysis using RMAT-20,
which is small enough to fit in memory and thus does not need
out-of-core operations, and RMAT-28, which requires out-of-core
operation in MR-MPI.  Table~\ref{table:prdetail} shows the amount of
time spent in each step of Figure~\ref{fig:pr2} for the MR-MPI
implementation and the equivalent Trilinos implementation.  For both
implementations, the matrix-vector multiplication (Step 2) is the most
costly step.  However, the MR-MPI algorithm is consistently slower
than the Trilinos implementation in each step.  This effect is more
pronounced for RMAT-28, where MR-MPI performs out-of-core operations.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline 
& \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
& MR-MPI & Trilinos & MR-MPI & Trilinos  \\
\hline
Step 1 (Compute adjustment) & 0.0203 & 0.00030 & 17.53 & 0.0838 \\
Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 & 341.26 & 2.9239 \\
Step 3 (Apply adjustment) & 0.0015 & 0.00015 & 2.14 & 0.0494 \\
Step 4 (Scale $y$)  & 0.0013 & 0.00014 & 1.74 & 0.0359 \\
Step 5 (Compute residual)  & 0.0207 & 0.00014 & 21.20 & 0.0457 \\
\hline
\end{tabular}
\caption{Detailed execution times (in seconds) per PageRank iteration
for each stage of the algorithm
in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
\label{table:prdetail}
\end{center}
\end{table}

To understand where MR-MPI spends its effort during the PageRank
computation, we break down the operation times further for the most
expensive Step 2 of the PageRank computation; the results are in
Table~\ref{table:prdetail2}.  In the table, we divide the {\it
collate()} operation into its two parts: an {\it aggregate()} that
involves parallel communication and a {\it convert()} that gathers
local key/value (KV) pairs with matching keys into a key-multivalue
(KMV) pair.  All interprocessor communication occurs in the {\it
aggregate()}.  While the time spent in {\it aggregate()} is
significant, it is small compared to the time spent in the two {\it
convert()} operations.  When out-of-core operations are required for
RMAT-28, the communication cost in the {\it aggregate()} is a smaller
percentage of total time for Step 2, as other data manipulation
operations such as {\it convert()} and {\it add()} require a
relatively larger percentage of the total time.  Further
experimentation is needed to determine whether reorganizing the
MapReduce data, as described in Section~\ref{subsec:graph_pagerank}.
would reduce the execution time of Step 2.  Organizing the data such
that both the matrix $A$ and vectors $x$ and $y$ are in the same
MapReduce object would eliminate the {\it copy()} and {\it add()}
functions, as well as the first {\it convert()}.  However, to update
the $y$ values at the completion of Step 2, either an additional {\it
convert()} or {\it aggregate()} with greater amounts of data would be
needed.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 2 & RMAT-20 & RMAT-28 \\
\hline
Copy $x$ to $y$                       &  1.1\% &  0.4\%\\
Add $A$ to $y$                        &  3.0\% & 14.3\% \\
Convert $y$                           & 23.5\% & 43.6\% \\
Reduce $y$                            &  7.7\% &  5.8\% \\
Aggregate $y$ (first part of collate) & 24.9\% &  6.8\% \\
Convert $y$ (second part of collate)  & 38.2\% & 27.4\% \\
Reduce $y$                            &  1.6\% &  1.7\% \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 2 of the MR-MPI 
PageRank algorithm in Figure~\ref{fig:pr2}.}
\label{table:prdetail2}
\end{center}
\end{table}

@@

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 5 & RMAT-20 & RMAT-28 \\
\hline
Add $y$ to $x$ &  2.2\% &  9.8\% \\
Convert $x$    & 95.1\% & 84.9\% \\
Reduce $x$     &  2.5\% &  5.3\% \\
MPI\_Allreduce &  0.2\% &  0.0\% \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 5 of
the MR-MPI PageRank algorithm in Figure~\ref{fig:pr2}.}
\label{table:prdetail5}
\end{center}
\end{table}

@@

For comparison, we wrote a parallel distributed-memory R-MAT
generation code with explicit MPI calls using a different
(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
must implement the parallelism explicitly whereas the codes of this
section relegate that task to the MR-MPI library.  A similar all-MPI
implementation of the PageRank algorithm required 1206 lines.  By
contrast an implementation of PageRank using the Trilinos parallel
linear algebra library (discussed in Section
\ref{subsec:results_pagerank}) required only 153 lines of code, since
an entire matrix-vector multiply can be invoked by a single call to
Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
required XXX lines of Java code.

>>>>

All data sets used R-MAT parameters $(a, b, c, d) = (0.57, 0.19, 0.19,
0.05)$ and generated eight edges per vertex (on average).  These
parameters create a highly skewed degree distribution, as indicated by
the maximum vertex degree in Table~\ref{t:rmats}.

@@

A more detailed analysis of the performance of the MR-MPI PageRank
algorithm is shown below.  We perform the analysis using RMAT-20,
which is small enough to fit in memory and thus does not need
out-of-core operations, and RMAT-28, which requires out-of-core
operation in MR-MPI.  Table~\ref{table:prdetail} shows the amount of
time spent in each step of Figure~\ref{fig:pr2} for the MR-MPI
implementation and the equivalent Trilinos implementation on 64 processors.  For both
implementations, the matrix-vector multiplication (Step 2) is the most
costly step.  However, the MR-MPI algorithm is consistently slower
than the Trilinos implementation in each step.  This effect is more
pronounced for RMAT-28, where MR-MPI performs out-of-core operations.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|r|r|}
\hline 
& \multicolumn{2}{|c|}{RMAT-20} & \multicolumn{2}{|c|}{RMAT-28} \\
& MR-MPI & Trilinos & MR-MPI & Trilinos  \\
\hline
Step 1 (Compute adjustment) & 0.0203 & 0.00030 & 17.53 & 0.0838 \\
Step 2 (Multiply $A^T x$) & 0.1377 & 0.00455 & 341.26 & 2.9239 \\
Step 3 (Apply adjustment) & 0.0015 & 0.00015 & 2.14 & 0.0494 \\
Step 4 (Scale $y$)  & 0.0013 & 0.00014 & 1.74 & 0.0359 \\
Step 5 (Compute residual)  & 0.0207 & 0.00014 & 21.20 & 0.0457 \\
\hline
\end{tabular}
\caption{Detailed execution times (in seconds) per PageRank iteration
on 64 processors
for each stage of the algorithm
in Figure~\ref{fig:pr2} using MR-MPI and Trilinos.}
\label{table:prdetail}
\end{center}
\end{table}

To understand where MR-MPI spends its effort during the PageRank
computation, we break down the operation times further for Step 2, the most
expensive step of the PageRank computation; the results are in
Table~\ref{table:prdetail2}.  In the table, we divide the {\it
collate()} operation into its two parts: an {\it aggregate()} that
involves parallel communication and a {\it convert()} that gathers
local key/value (KV) pairs with matching keys into a key-multivalue
(KMV) pair.  All interprocessor communication occurs in the {\it
aggregate()}.  While the time spent in {\it aggregate()} is
significant, it is small compared to the time spent in the two {\it
convert()} operations.  When out-of-core operations are required for
RMAT-28, the communication cost in the {\it aggregate()} is a smaller
percentage of total time for Step 2, as other data manipulation
operations such as {\it convert()} and {\it add()} require a
relatively larger percentage of the total time.  Further
experimentation is needed to determine whether reorganizing the
MapReduce data, as described in Section~\ref{subsec:graph_pagerank},
would reduce the execution time of Step 2.  Organizing the data such
that both the matrix $A$ and vectors $x$ and $y$ are in the same
MapReduce object would eliminate the {\it copy()} and {\it add()}
functions, as well as the first {\it convert()}.  However, to update
the $y$ values at the completion of Step 2, either an additional {\it
convert()} or {\it aggregate()} with greater amounts of data would be
needed.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 2 & RMAT-20 & RMAT-28 \\
\hline
Copy $x$ to $y$                       &  1.1\% &  0.4\%\\
Add $A$ to $y$                        &  3.0\% & 14.3\% \\
Convert $y$                           & 23.5\% & 43.6\% \\
Reduce $y$                            &  7.7\% &  5.8\% \\
Aggregate $y$ (first part of collate) & 24.9\% &  6.8\% \\
Convert $y$ (second part of collate)  & 38.2\% & 27.4\% \\
Reduce $y$                            &  1.6\% &  1.7\% \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 2 of the MR-MPI 
PageRank algorithm in Figure~\ref{fig:pr2}; experiments used 64 processors.}
\label{table:prdetail2}
\end{center}
\end{table}

@@

\begin{table}[htb]
\begin{center}
\begin{tabular}{|l|r|r|}
\hline 
PageRank Step 5 & RMAT-20 & RMAT-28 \\
\hline
Add $y$ to $x$ &  2.2\% &  9.8\% \\
Convert $x$    & 95.1\% & 84.9\% \\
Reduce $x$     &  2.5\% &  5.3\% \\
MPI\_Allreduce &  0.2\% &  0.0\% \\
\hline
\end{tabular}
\caption{Percentage of execution time in each operation of Step 5 of
the MR-MPI PageRank algorithm in Figure~\ref{fig:pr2}; experiments
used 64 processors.}
\label{table:prdetail5}
\end{center}
\end{table}

@@

For comparison, we wrote a parallel distributed-memory R-MAT
generation code with explicit MPI calls using a different
(non-MapReduce) algorithm than that of Figures \ref{fig:rmat} and
\ref{fig:rmat2}.  It required 1006 lines of code, mainly because it
must implement the parallelism explicitly, whereas the codes of this
section relegate that task to the MR-MPI library.  A similar all-MPI
implementation of the PageRank algorithm required 1206 lines.  By
contrast an implementation of PageRank using the Trilinos parallel
linear algebra library (discussed in Section
\ref{subsec:results_pagerank}) required only 153 lines of code, since
an entire matrix-vector multiplication can be invoked by a single call to
Trilinos.  Our collaborators (see Acknowledgments) have a Hadoop
implementation of the SSSP algorithm of Figure \ref{fig:sssp}.  It
required 697 lines of Java code.
